---
title: "Glasshouse tomato production"
description: |
  Description
last-modified: "`r Sys.Date()`"
bibliography: ref.bib
---


This example is inspired from @meadStatisticalPrinciplesDesign2012 page 475.

::: callout-note

- The aim is to study the production of glasshouse tomatoes under different air and soil temperatures.  
- There are 8 glasshouse compartments.
- Each compartment contained two large troughs where tomatoes are grown.
- For each trough, half of the soil would be heated or unheated.
- The compartment was kept at minimum either 13°C or 16°C.
- Heated soil was kept at either 18°C or 21°C.

:::


```{r}
#| message: false
library(edibble)
tomato_units <- set_units(compartment = 8,
                               trough = nested_in(compartment, 2),
                           halftrough = nested_in(trough, 2))

tomato_trts <- set_trts(air_temp = c("13°C", "16°C"),
                          heated = c("Yes", "No"),
                       soil_temp = conditioned_on(heated,
                                          "Yes" ~ c("18°C", "21°C"),
                                           "No" ~ "Natural"))

tomato_design <- (tomato_units + tomato_trts) %>% 
  allot_table(air_temp ~ compartment,
                heated ~ halftrough,
             soil_temp ~ halftrough,
     # algorithms for each allotment
     # Note: blocksdesign is from the `blocksdesign` package
                              order = c("random", 
                                        "random",
                                        "blocksdesign"),
                              label_nested = c(halftrough, trough), 
                              seed = 1)
```


```{r}
#| code-fold: true
library(ggplot2)
tomato_design %>% 
  ggplot(aes(trough, halftrough)) +
  geom_rect(aes(fill = air_temp), xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) + 
  scale_fill_manual(values = c("#b5cbde", "#b3d5a4")) +
  ggnewscale::new_scale_fill() +
  geom_tile(aes(fill = soil_temp), color = "black", width = 0.9) +
  scale_fill_manual(values = c("#e57f7f", "#7a0000", "#8c8c8c"), na.value = "#8c8c8c") +
  facet_wrap(~compartment, ncol = 2) +
  theme(axis.text = element_text(color = "black", margin = margin(t = 5, r = 5)))
```

